opXRD: Open Experimental Powder X-ray Diffraction Database

### 2025-03-01 chapters/04_our_dataset.tex
Commited Feburary changes

In collaboration with several other institutions we have collected a dataset of diffractograms stemming from experiments, some of them labeled with corresponding structural information. The following paragraphs detail the contribution of each research group. Most data was collected using \ce{Cu} radiaton sources which have a  $K_{\alpha1}$ wavelength of $\lambda=1.54056\text{\AA}$ and a $K_{\alpha2}$ wavelength of $\lambda=1.54439\text{\AA}$. We filtered the submitted datasets to exclude patterns with invalid features such as the pattern containing only one unique recorded angle or negative angles, the pattern containing less than 50 recorded angles total, or all intensities being zero. The below Table~$\eqref{tab:merged}$ provides an overview of the opXRD database.

\begin{table}[!htb]
\centering
\caption{\footnotesize The opXRD database: The availability of the chemical composition, spacegroups, lattice parameters, and atomic coordinates of the underlying samples are indicated by the columns ``Comp.'', ``Spg.'', ``Lattice'' and ``Atom coords.'' respectively.}
\label{tab:merged}
\scalebox{0.75}{
\begin{adjustbox}{center}
\begin{tabular}{@{}lllllll@{}}
%\begin{tabular}{@{}llccccl@{}}
\toprule
\textbf{Institution} & \textbf{No. patterns} & \ \textbf{Comp.} & \ \ \textbf{Spg.} & \textbf{Lattice} & \textbf{Atom coords.} & \textbf{Research Project} \\
\midrule
CNRS        &  1052          & \ \cmark   & \partialcheck{85}   & \cmark   &  \ \ \quad \partialcheck{85}   & Diffraction data extracted from the COD \\ 
HKUST(GZ)   &  21+499          & \ \cmark   & \cmark   & \cmark   & \ \ \quad \cmark   & Phase/structure identification dataset  \\ 
USC         &  338          & \ \cmark   & \cmark   & \partialcheck{90}   & \ \ \quad \xmark   & Study of CuNi and CuAl alloys          \\ 
EMPA        &  790          &\ \cmark   & \partialcheck{63}   & \xmark   & \ \ \quad \xmark   & Tin halide perovskites, Zn-V-N libraries \\ 
INT         &  19,780        &\ \xmark   & \xmark   & \xmark   & \ \ \quad \xmark   & Compilation of various projects        \\ 
IKFT        &  64           &\ \xmark   & \xmark   & \xmark   & \ \ \quad \xmark   & Commercial catalysts, bulk materials   \\ 
LBNL        &  n/a          &\ \xmark   & \xmark   & \xmark   & \ \ \quad \xmark   & Perovskites precursors, Mn-Sb-O system  \\ 
\bottomrule
\end{tabular}
\end{adjustbox}}
\end{table}

The variance of the data was analyzed using a dimensional reduction technique called principal component analysis (PCA). PCA can be applied to datasets $X \subset \mathbb{R}^N$ to reduce the number of components needed to describe points $p \in X$ up to some tolerance in lost accuracy. In the context of PCA, the cumulative explained variance ratio is a measure of how much of the variance in the dataset $X$ can be explained using a given number of components. For example, if performing PCA on a dataset $X \subset \mathbb{R}^{10}$ yields a cumulative variance ratio of 99\%, then each point in the dataset can be represented using only 5 degrees of freedom up to a minor loss in accuracy. For a rigorous definition of PCA and the explained variance ratio, we refer to the literature \cite{Jolliffe2016}. \\

Here, PCA was performed on datasets of X-ray diffraction patterns. These datasets $X$ are subsets of $\mathbb{R}^{N}$ with $N=N_{\text{values}} =512$ since each pattern $p \in X$ was standardized to have 512 intensity values spread out evenly from $0\degree$ to $180\degree$ using zero padding and interpolation with cubic splines. Hence the maximal components that could be needed to describe a dataset of diffraction data in this context is $N_{\text{values}}=512$. However, the maximal number of components is even lower for datasets that contain less than 512 patterns. In this case, the maximal number of components is equal to the number of patterns in the dataset since each pattern can add at most one degree of freedom to the dataset $X \in \mathbb{R}^N$. Hence the maximum number of components $N_{\text{max}}$ of a pattern dataset $X$ is given as follows:
\begin{align}
    N_{\text{max}} = \min(N_{\text{values}}, N_{\text{patterns}}).
\end{align}
Here $N_{\text{values}} = N =512$ is the number of recorded intensity values per pattern and $N_{\text{patterns}}$ is the number of patterns in the dataset $X$. Figure~$\eqref{fig:components}$ below shows the cumulative explained variance ratio over the fraction of maximal No. components $N_{\text{max}}$ as defined above. In this figure, a faster convergence of the cumulative variance ratio towards one indicates that the patterns in this data are relatively similar. \\

The degree of variation between the patterns is different for each contribution. For example, the CRNS and the HKUST contributions each are collections that encompass many research projects over a large period of time and hence exhibit a high degree of variability between individual patterns. In contrast, the contributions by the USC and the LBNL contain many very similar patterns. The patterns in the USC datasets are similar because the underlying samples are all variations of \ce{CuNi} and \ce{CuAl} alloys collected on the same machine. The patterns submitted by the LBNL are similar because they stem from in-situ recordings where several hundred or several thousand patterns were collected over time per sample while they were undergoing physical processes.  
\pagebreak

These effects can be observed in figure~$\eqref{fig:components}$: The cumulative explained variance ratios of the datasets contributed by the USC and LBNL converge towards one significantly faster than the datasets contributed by the CNRS and HKUST. \\

\begin{figure}[!htb]
    \centering
    \includegraphics[width=0.85\linewidth]{figures/component_fractions.png}
    \caption{Explained variance ratio over the fraction of the maximum number of components for each dataset contributed to opXRD. Each dataset is named after the contributing institutions. Datasets contributed by the same institution are labeled alphabetically in the order in which they are described in the texts towards the end of this section.}
    \label{fig:components}
\end{figure}

Figure~$\eqref{fig:histograms}$ below provides an overview of the distributions of pattern and structure properties in the opXRD dataset. Nearly all patterns have an angular resolution smaller than $\Delta(2\theta) = 0.1 ^\circ$. Here the angular resolution is defined as the range of recorded angles divided by the number of recorded intensity values along that range. For most patterns, the lowest recorded angle is smaller than $30 ^\circ$ and the highest recorded angle is smaller than $120 ^\circ$.

@@

The following texts describe the datasets contributed by each collaborating institution. Each text includes a description of the investigated materials and how X-ray diffraction data was collected.

@@

A small-scale experimental powder X-ray database called the X-Ray Phase Identification Public Experimental Dataset (XRed) was established over the past two years.\footnote{XRed can be accessed on GitHub at \url{https://github.com/WPEM/XRED}.} The primary objective of this database is to support the development of intelligent phase identification technology, serving as a foundation for data collection for future large-scale machine learning calculations. The XRed dataset primarily focuses on metal and metal-oxide particles, with data generated using instruments such as the Empyrean 3.0, Aeris, and Bruker D8 Advance diffractometers, all utilizing \ce{Cu} X-ray sources. Hundreds of pXRD patterns have been obtained through XRD refinement, enabling the determination of detailed lattice constants, atomic sizes, and other structural parameters. These patterns are labeled with CIF files that document the refined structures. The dataset is organized by elemental systems and includes original experimental files ranging from single-phase to five-phase mixtures, as well as mixtures studied for different tasks. \\

Another experimental dataset integrated with opXRD is a collection of powder diffraction data sourced from open-access publications and collaborating institutions. All collaborating institutions provided this data with full authorization for research purposes. This dataset provides broader coverage of chemical elements and includes ionic crystals, atomic crystals, and metallic crystals. This dataset surpasses Xred in size, containing 499 entries. However, contrary to Xred these data do not come paired with corresponding crystal structures. \\

Xqueryer,\footnote{Xqueryer can be accessed at \url{https://xqueryer.caobin.asia/}.} a recently launched cutting-edge free online platform for intelligent structure identification, will also serve as a source of data for the opXRD database. Users can upload experimental pXRD data to identify crystal structures and extract all relevant information available in the Materials Project database\cite{jain2013commentary}. Uploaded data are securely stored and integrated into the comprehensive opXRD database to progress toward a large-scale, long-term experimental resource. All usage complies with Xqueryer's terms of service.

@@

X-ray diffraction data was collected from a wide range of research projects conducted at the Institute of Nanotechnology over the past years. A major part of the research focused on high-entropy materials, which involved incorporating many different elements into single-phase structures, leading to peak shifts or phase separations. Most of those multi-component complex materials appeared in various structures, including rock-salt, spinel, fluorite, perovskite, and delafossite. The samples were prepared either in powder or in bulk form; therefore, powder XRD was performed on samples with adjusted height. The samples were prepared using various synthesis techniques, mostly solid-state or wet chemical syntheses, to obtain the desired structures. Consequently, particle size and crystallinity varied significantly. The sample set also includes samples that were not successfully measured or where phases could not be identified. \\

@@

In total three different datasets were collected. The first and second datasets focused on metal-halide perovskite precursor solutions. The first dataset was collected from spin-coating triple-cation metal-halide perovskite precursor solutions with the composition \ce{Cs}$_{0.05}$(MA$_{0.23}$FA$_{0.77}$)\ce{Pb}$_{1.1}$(\ce{I}$_{0.77}$\ce{Br}$_{0.23}$)$_{3}$ onto various substrates. Here, MA stands for Methylammonium and FA stands for Formamidinium. The substrates onto which these solutions were coated include glass, which is amorphous, and $\ce{GaAs}$ wafers, which are single crystalline. Other substrates were stacks of glass/indium tin oxide, stacks of \ce{GaAs}/CIGS, and stacks of glass/CIGS. Here, CIGS stands for a stack of \ce{Mo}, \ce{Cu}(\ce{In}, \ce{Ga})\ce{Se2},\ce{Cds} and \ce{ZnO}. Some of the substrates were additionally covered with a self-assembling monolayer of MeO-2PACz. The $\ce{GaAs}$ substrates were prepared by Dr. Jiro Nishinaga from the National Institute of Advanced Industrial Science and Technology (AIST) in Japan and the glass/CIGS substrates by Dr. Christian Kaufmann and his team at Helmholtz-Zentrum Berlin (HZB) in Germany. Data collection was performed in situ during thin-film deposition using a custom-made spin-coating and annealing stage. \\

A second dataset was collected from spin-coating metal-halide perovskite precursor solutions with varying compositions of \ce{MAPb(I_{1-x}Br_x)3} spin-coated onto glass substrates. Here, MA = Methylammonium and x = 0, 0.33, 0.5, 0.67, 1. The substrates were preheated to different temperatures including $30\degree $C, $50\degree $C, $70\degree $C, and $90\degree $C, and the spin-coating process was performed at a constant temperature on the preheated substrates. For both datasets, diffraction data were continuously measured during spin-coating, chemical induction of crystallization, and annealing of the samples, at $100\degree$C and $110\degree$C respectively. The diffraction data was recorded with a frequency of about $0.56 
 \ 1/\si{s}$ and $0.54 \ 1/\si{s}$. Each in situ measurement consisted of about 500 to 1000 individual diffractograms. Depending on the substrate, each series of diffractograms shows an evolution from substrate only to a combination of polycrystalline perovskite, $\ce{PbI2}$ and substrate via several intermediate phases. \\

@@

A third dataset was collected by observing the phase evolution of an Mn-Sb-O system with varying annealing temperatures. The temperatures used to analyze the crystal structure of the Mn-Sb-O system were chosen depending on the number of phase transitions appearing for a certain temperature range. Few changes in the crystal structure appear between room temperature and 300$\degree$C and phase transitions appeared from 300$\degree$C until 850$\degree$C. No phase transition appeared when cooling down. Therefore, the crystal structure was measured every 100$\degree$C between room temperature and 300$\degree$C; every 50$\degree$C between 300$\degree$C and 850$\degree$C; and every 200$\degree$C when cooling down. The heating and cooling rates are fixed for all the experiments at 50$\degree$C/min and the holding time is fixed at 2 min. \\

The data was collected using the \textit{in situ}  Rigaku-SmartLab3kW diffractometer. This tool operates with SmartLab Studio II software, which can measure the X-ray diffraction during the annealing process. This enables directly showing all the phase transitions when annealing in various atmospheres such as \ce{O2}, \ce{Ar}, and \ce{NH3}. Phase transitions are analyzed with the \textit{in situ} XRD tool up to 850$\degree$C in this work. Most of the \textit{in situ} experiments were performed under an air-like 20\% \ce{O2} and 80\% \ce{Ar} environment is chosen (\ce{Ar} flow: 50 \si{sccm}, \ce{O2} flow: 10 \si{sccm}). When a 100\% Ar environment is fixed, an Ar flow of 60 \si{sccm} is input. The Bragg-Brentano (BB) mode is preferred in terms of geometry because it is more adapted in the analysis of scarce phases such as  \ce{MnSb2O6} rutile. The angular step used in the recording was 0.01$\degree$ and the scanning rate was 10$\degree$/min.

@@

>>>>

In collaboration with several other research institutions, we have collected a database of $\numpatterns$ experimental patterns of which 2179 are at least partially labeled with structural information of the underlying material. The following research institutions contributed data to the opXRD database: Paris Sciences et Lettres University University (PSL), Hong Kong University of Science and Technology (Guangzhou) (HKUST), University of Southern California (USC), Lawrence Berkeley National Laboratory (LBNL), Empa–Swiss Federal Laboratories for Materials Science and Technology (EMPA) and the Karlsruhe Institute of Technology (KIT). \newline

We filtered the submitted datasets to exclude patterns with invalid features such only one unique recorded angle, negative angles, less than 50 recorded angles total, or all intensities being zero. Table~$\eqref{tab:merged}$ provides an overview of the opXRD database.

\begin{table}[!htb]
\centering
\caption{\footnotesize Overview of the contributions to the opXRD database: The availability of the chemical composition, spacegroups, lattice parameters, and atomic coordinates of the underlying samples are indicated by the columns ``Comp.'', ``Spg.'', ``Lattice'' and ``Atom coords.'' respectively.}
\label{tab:merged}
\scalebox{0.75}{
\begin{adjustbox}{center}
\begin{tabular}{@{}lllllll@{}}
%\begin{tabular}{@{}llccccl@{}}
\toprule
\textbf{Institution} & \textbf{No. patterns} & \ \textbf{Comp.} & \ \ \textbf{Spg.} & \textbf{Lattice} & \textbf{Atom coords.} & \textbf{Research Project} \\
\midrule
CNRS        &  1052          & \ \cmark   & \partialcheck{85}   & \cmark   &  \ \ \quad \partialcheck{85}   & Diffraction data extracted from the COD \\ 
HKUST(GZ)   &  520          & \ \cmark   & \cmark   & \cmark   & \ \ \quad \cmark   & Phase identification dataset  \\ 
USC         &  338          & \ \cmark   & \cmark   & \partialcheck{90}   & \ \ \quad \xmark   & Study of CuNi and CuAl alloys          \\ 
EMPA        &  770          &\ \cmark   & \partialcheck{63}   & \xmark   & \ \ \quad \xmark   & Metal halide perovskites, Zn-V-N libraries \\ 
INT         &  19,796        &\ \xmark   & \xmark   & \xmark   & \ \ \quad \xmark   & Compilation of various projects        \\ 
IKFT        &  64           &\ \xmark   & \xmark   & \xmark   & \ \ \quad \xmark   & Commercial catalysts, metals, metal oxides   \\ 
LBNL        &  70,012          &\ \xmark   & \xmark   & \xmark   & \ \ \quad \xmark   & Perovskites precursors, Mn-Sb-O system  \\ 
\bottomrule
\end{tabular}
\end{adjustbox}}
\end{table}

The variance of the data was analyzed using a principal component analysis (PCA). PCA can be applied to datasets $X \subset \mathbb{R}^N$ to reduce the number of components needed to describe points $p \in X$ up to some tolerance in lost accuracy. In the context of PCA, the cumulative explained variance ratio is a measure of how much of the variance in the dataset $X$ can be explained using a given number of components. For a rigorous definition of PCA and the explained variance ratio, we refer to the literature \cite{Jolliffe2016}. \newline

Here, PCA was performed on datasets of X-ray diffraction patterns. These datasets $X$ are subsets of $\mathbb{R}^{N}$ with $N=512$ since each pattern $p \in X$ was standardized to have 512 intensity values spread out evenly from $0\degree$ to $180\degree$ using zero padding and interpolation with cubic splines. Hence the maximal components that could be needed to describe a dataset of diffraction data in this context is $N=512$. However, the maximal number of components is even lower for datasets that contain less than 512 patterns. In this case, the maximal number of components is equal to the number of patterns in the dataset since each pattern can add at most one degree of freedom to the dataset $X \in \mathbb{R}^N$. Hence the maximum number of components $N_{\text{max}}$ of a pattern dataset $X$ is given as follows:
\begin{align}
    N_{\text{max}} = \min(N_{\text{values}}, N_{\text{patterns}}).
      \label{eq:nmax}
\end{align}
Here $N_{\text{values}} = N =512$ is the number of recorded intensity values per pattern and $N_{\text{patterns}}$ is the number of patterns in the dataset $X$. Figure~$\eqref{fig:components}$ below shows the cumulative explained variance ratio over the fraction of maximal No. components $N_{\text{max}}$ as defined above. In this figure, a faster convergence of the cumulative variance ratio towards one indicates that the patterns in this data are relatively similar. \\

The degree of variation between the patterns is different for each contribution. For example, the CRNS and the HKUST contributions each are collections that encompass many research projects over a large period of time and hence exhibit a high degree of variability between individual patterns. In contrast, the contributions by USC and LBNL contain many very similar patterns. The patterns in the USC datasets are similar because the underlying samples are all variations of \ce{CuNi} and \ce{CuAl} alloys collected on the same machine. The patterns submitted by LBNL are similar because they stem from in-situ recordings where several hundred or several thousand patterns were collected over time per sample while they were undergoing physical conversion processes.  
\pagebreak

These effects can be observed in Figure~$\eqref{fig:components}$: The cumulative explained variance ratios of the datasets contributed by USC and LBNL converge towards one significantly faster than the datasets contributed by CNRS and HKUST. \\

\begin{figure}[!htb]
    \centering
    \includegraphics[width=0.85\linewidth]{figures/component_fractions.png}
    \caption{Explained variance ratio over the fraction of the maximum number of components for each dataset contributed to the opXRD database. Here the maximal No. components refers to $N_{\text{max}}$ as defined in equation $\eqref{eq:nmax}$. Each dataset is named after the contributing institutions. Datasets contributed by the same institution are labeled alphabetically in the order in which they are described in the texts towards the end of this section.}
    \label{fig:components}
\end{figure}

Figure~$\eqref{fig:histograms}$ provides an overview of the distributions of pattern and structure properties in the opXRD dataset. Nearly all patterns have an angular resolution smaller than $\Delta(2\theta) = 0.1 ^\circ$. Here the angular resolution is defined as the range of recorded angles divided by the number of recorded intensity values along that range. For most patterns, the lowest recorded angle is smaller than $30 ^\circ$ and the highest recorded angle is smaller than $120 ^\circ$. The start-to-end angle distribution reveals that all diffractograms start in a narrow window between $0\degree$ and approximately $50\degree$, while they end between $50\degree$ and $150\degree$, with the majority of patterns going from $0\degree$ to approximately $70\degree$. This means that, unlike in most ML approaches using synthetic data over the full angle range with fixed resolution, the opXRD dataset has a strongly varying angle range and resolution, which requires additional pre-processing methods such as padding and interpolation, or more flexible ML models beyond standard CNNs.

@@

In the following, we will describe the datasets contributed by each of the collaborating research groups and institutions. Each paragraph includes a description of the investigated materials and how X-ray diffraction data was collected. Most data was collected using \ce{Cu} radiaton sources which has a $K_{\alpha1}$ wavelength of $\lambda=1.54056\text{\AA}$ and a $K_{\alpha2}$ wavelength of $\lambda=1.54439\text{\AA}$.

@@

A small-scale experimental powder X-ray database, the X-Ray Phase Identification Public Experimental Dataset (XRed), has been developed over the past two years (\url{https://github.com/WPEM/XRED}). The primary goal of XRed is to support the advancement of intelligent phase identification technology by providing a foundation for data collection in future large-scale machine learning applications. XRed primarily focuses on metal and metal-oxide particles, with data collected using diffractometers such as the Empyrean 3.0, Aeris, and Bruker D8 Advance, all employing \ce{Cu} X-ray sources. The dataset contains hundreds of pXRD patterns, obtained through XRD refinement, which enable the determination of lattice constants, atomic sizes, and other structural parameters. Each pattern is labeled with a corresponding CIF that documents the refined structure. Data are categorized by elemental systems and include original experimental files, spanning single-phase to five-phase mixtures, as well as mixtures designed for various research tasks. \\

In addition to XRed, the opXRD database integrates an experimental dataset composed of powder diffraction data sourced from open-access publications and collaborating institutions. These institutions have provided the data with full authorization for research purposes. Compared to XRed, this dataset offers broader chemical element coverage, encompassing ionic, atomic, and metallic crystals. It is also larger, containing 499 entries. However, unlike XRed, these data entries are not accompanied by CIF files. \\

Another data source for opXRD is Xqueryer (\url{https://xqueryer.caobin.asia/}), a recently launched free online platform for intelligent structure identification. Users can upload experimental pXRD data to identify crystal structures and extract relevant information from the Materials Project database\cite{jain2013commentary}. Uploaded data are securely stored and integrated into the opXRD database, contributing to its development as a large-scale, long-term experimental resource. All data usage adheres to Xqueryer's terms of service.

@@

X-ray diffraction data was collected from a wide range of research projects conducted at the Institute of Nanotechnology over the past 10 years. A major part of the research focused on high-entropy materials, which involved incorporating many different elements into single-phase structures, leading to peak shifts or phase separations. Most of those multi-component complex materials appeared in various structures, including rock-salt, spinel, fluorite, perovskite, and delafossite. The samples were prepared either in powder or in bulk form; therefore, powder XRD was performed on samples with adjusted height. The samples were prepared using various synthesis techniques, mostly solid-state or wet chemical syntheses, to obtain the desired structures. Consequently, particle size and crystallinity varied significantly. The sample set also includes samples that were not successfully measured or where phases could not be identified. \\

@@

In total four different datasets were collected. The first and second datasets focused on metal-halide perovskite thin-films, (...). \\

The first dataset (LBNL-A) was collected from spin-coating and annealing triple-cation metal-halide perovskite precursor solutions with the composition \ce{Cs}$_{0.05}$(MA$_{0.23}$FA$_{0.77}$)\ce{Pb}$_{1.1}$(\ce{I}$_{0.77}$\ce{Br}$_{0.23}$)$_{3}$ onto various substrates. Here, MA stands for Methylammonium and FA stands for Formamidinium. The substrates onto which these solutions were coated include glass, which is amorphous, and $\ce{GaAs}$ wafers, which are single crystalline. Other substrates were stacks of glass/indium tin oxide, stacks of \ce{GaAs}/CIGS, and stacks of glass/CIGS. Here, CIGS stands for a stack of \ce{Mo}, \ce{Cu}(\ce{In}, \ce{Ga})\ce{Se2},\ce{Cds} and \ce{ZnO}. Some of the substrates were additionally covered with a self-assembling monolayer of MeO-2PACz. The $\ce{GaAs}$ substrates were prepared by Dr. Jiro Nishinaga from the National Institute of Advanced Industrial Science and Technology (AIST) in Japan and the glass/CIGS substrates by Dr. Christian Kaufmann and his team at Helmholtz-Zentrum Berlin (HZB) in Germany. Data collection was performed in situ during thin-film deposition using a custom-made spin-coating and annealing stage \cite{song2019}.

A second dataset (LBNL-B) was collected from spin-coating metal-halide perovskite precursor solutions with varying compositions of \ce{MAPb(I_{1-x}Br_x)3} spin-coated onto glass substrates. Here, MA = Methylammonium and x = 0, 0.33, 0.5, 0.67, 1. The substrates were preheated to different temperatures including $30\degree $C, $50\degree $C, $70\degree $C, and $90\degree $C, and the spin-coating process was performed at a constant temperature on the preheated substrates. For both datasets, diffraction data were continuously measured during spin-coating, chemical induction of crystallization, and annealing of the samples, at $100\degree$C and $110\degree$C respectively. The diffraction data was recorded with a frequency of about $0.56 
 \ 1/\si{s}$ and $0.54 \ 1/\si{s}$. Each in situ measurement consisted of about 500 to 1000 individual diffractograms. Depending on the substrate, each series of diffractograms shows an evolution from substrate only to a combination of polycrystalline perovskite, $\ce{PbI2}$ and substrate via several intermediate phases. \\

@@

A third dataset (LBNL-C) was collected by observing the phase evolution of an Mn-Sb-O system with varying annealing temperatures. The temperatures used to analyze the crystal structure of the Mn-Sb-O system were chosen depending on the number of phase transitions appearing for a certain temperature range. Few changes in the crystal structure appear between room temperature and 300$\degree$C and phase transitions appeared from 300$\degree$C until 850$\degree$C. No phase transition appeared when cooling down. Therefore, the crystal structure was measured every 100$\degree$C between room temperature and 300$\degree$C; every 50$\degree$C between 300$\degree$C and 850$\degree$C; and every 200$\degree$C when cooling down. The heating and cooling rates were fixed for all the experiments at 50$\degree$C/min and the holding time was fixed to 2 min. \\

This data was collected using the \textit{in situ}  Rigaku-SmartLab3kW diffractometer. This tool operates with SmartLab Studio II software, which can measure the X-ray diffraction during the annealing process. This enables directly showing all the phase transitions when annealing in various atmospheres such as \ce{O2}, \ce{Ar}, and \ce{NH3}. Phase transitions are analyzed with the \textit{in situ} XRD tool up to 850$\degree$C in this work. Most of the \textit{in situ} experiments were performed under an air-like 20\% \ce{O2} and 80\% \ce{Ar} environment is chosen (\ce{Ar} flow: 50 \si{sccm}, \ce{O2} flow: 10 \si{sccm}). When a 100\% Ar environment is fixed, an Ar flow of 60 \si{sccm} is input. The Bragg-Brentano (BB) mode is preferred in terms of geometry because it is more adapted in the analysis of scarce phases such as  \ce{MnSb2O6} rutile. The angular step used in the recording was 0.01$\degree$ and the scanning rate was 10$\degree$/min.

@@

The fourth and final dataset (LBNL-D) was collected from (...). To prepare these compounds thiol-functionalized UiO-66-type Zr-based MOFs were added as additives to the PbI2 precursor solution followed by a two-step synthesis method to form FAMAPb(I,Br)3 films. XRD data was collected (...).

### 2025-03-01 chapters/05_how_to_use.tex
Added alphabetic labeling rule in Usage section
Commited Feburary changes

The opXRD database is hosted on Zenodo\footnote{The dataset can be found at \url{https://zenodo.org/records/14278656}.} and can be downloaded by any user without any barriers or restrictions. The database comes in two files, ``opxrd.zip'' and ``opxrd\_in\_situ.zip''. The latter contains the in-situ data which contains highly correlated patterns recorded through time series measurements. Patterns belonging to time series measurements have filenames that indicate the measurement series they belong to as well as their order in that series. \\

Next to the availability of the opXRD dataset on Zenodo, we also provide a Python library ``opxrd'' to easily download and interface with the dataset. The instructions for how to install this library can be found in the repository associated with the library.\footnote{The repository to this library can is located at \url{https://github.com/aimat-lab/opxrd}.} The opxrd library allows the dataset to be accessed through one simple command: \pyth{OpXRD.load(root_dirpath)}. If the database is locally available under \pyth{root_dirpath} this command loads the library from this location. If the database is not available locally at this location, the database is automatically downloaded to \pyth{root_dirpath}. \\

Our library further includes options for standardization, plotting, and the conversion to \emph{PyTorch} tensors. We provide a Jupyter Notebook\footnote{This notebook can be found here: \url{https://colab.research.google.com/github/aimat-lab/opXRD/blob/main/opxrd/usage.ipynb}.} that showcases these functionalities in more detail. This notebook also illustrates how to interface with the OpXRD database through Python.

>>>>

The opXRD database is hosted on Zenodo (\url{https://zenodo.org/records/14278656}) and can be downloaded by any user without any barriers or restrictions.
\\

Next to the availability of the opXRD dataset on Zenodo, we also provide a Python library ``opxrd'' to easily download and interface with the dataset. The instructions for how to install this library can be found in the repository associated with the library. The repository to this library is located at \url{https://github.com/aimat-lab/opxrd}. \\

The opxrd library includes options for data-loading, for standardization, plotting, and the conversion to \emph{PyTorch} tensors. We provide a Jupyter Notebook %\footnote{This notebook can be found here: \url{https://colab.research.google.com/github/aimat-lab/opXRD/blob/main/opxrd/usage.ipynb}.}
that showcases these functionalities in more detail. This notebook also illustrates how to interface with the opXRD database through Python.

### 2025-03-01 chapters/06_summary_and_outlook.tex
Updated introductory paragraphs on rietveld refinement
Merged Xqueyer paragraph into contribution entry
Commited table version with footnote directly underneath
Added histograms, PCA plots + paragraph on notebook; Added labeled column in existing datasets;
Updated research projects in overview table
Commited Feburary changes

With the opXRD dataset, a curation of XXX unlabeled and YYY labeled experimental powder XRD patterns from a wide range of different materials systems, we provide the largest currently available source of experimental XRD patterns. With this, we address the need for experimental data that arises when developing algorithms and analysis tools for XRD data, both based on machine learning and classical approaches. The data can be used for the actual method development and for testing. Our dataset is a valuable and so far missing resource to drive further developments in the automated analysis of XRD data.

\paragraph{How to contribute}

Rather than a finished project, the opXRD dataset is an ongoing effort to collect experimental powder XRD data. We invite everyone in possession of experimental powder XRD data to submit it to our dataset, to further improve the dataset and aid further developments in this field. Our submission page (\url{https://xrd.aimat.science/}) and submission helper software will be kept available to collect more data. We will keep updating and maintaining the dataset with incoming new submissions.

%We have launched Xqueryer (\url{http://xqueryer.caobin.asia/}), a cutting-edge, free online platform for intelligent structure identification. Users can upload experimental pXRD data to identify crystal structures and extract all relevant information available in the Materials Project database. All usage complies with Xqueryer's terms of service. Uploaded data are securely stored and integrated into the opXRD database.

>>>>

With the opXRD dataset, a curation of $\numpatterns$ unlabeled and 2179 at least partially labeled experimental powder XRD patterns from a wide range of different materials systems, we provide the largest currently available source of experimental XRD patterns. With this, we address the need for experimental data that arises when developing algorithms and analysis tools for XRD data, both based on machine learning and classical approaches. The data can be used for the actual method development and for testing. Our dataset is a valuable and so far missing resource to drive further developments in the automated analysis of XRD data.

% How to contribute
Rather than a finished project, the opXRD dataset is an ongoing effort to collect experimental powder XRD data. We invite everyone who is working in the area of experimental powder XRD to submit it to the dataset, in order to further improve the utility of the dataset and thus aid further developments in this field. Our submission page (\url{https://xrd.aimat.science/}) and submission helper software will be kept available to collect more data. We will keep updating and maintaining the dataset with new incoming submissions.
